---
title: "region_analysis_Pacific_Mountain"
author: "Li Li"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Utility functions

```{r}

library(dplyr)
library(MASS)
library(faraway)
library(lmtest)

# calculate RMSE
rmse = function(actual, predicted) {
  
  sqrt(mean((actual - predicted) ^ 2))
  
}


# calculate leave one out cross validation
calc_loocv_rmse = function(model) {
  
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
  
}


library(readr)

cars_data = read_csv("used_truck_data_with_region.csv")

cars_data$region = as.factor(cars_data$region)

cars_data$state = as.factor(cars_data$state)




cars_pacific = subset(cars_data, cars_data$region == "Pacific")

cars_mountain = subset(cars_data, cars_data$region == "Mountain")







```




## Pacific Region


```{r}

set.seed(420420420)

size_num = 5000

cars_idx = sample(nrow(cars_pacific), size_num)

cars_trn = cars_pacific[cars_idx, ]

cars_tst = cars_pacific[-cars_idx, ]


fit_common = lm(price ~ engine_displacement + engine_cylinders + highway_fuel_economy + horsepower + maximum_seating + transmission + 
                  mileage + year + wheel_system + power_hp + bed_length +
                  cabin + salvage + franchise_dealer + state, data = cars_trn)


summary(fit_common)


n = length(resid(fit_common))

# Apple BIC
fit_reduced = step(fit_common, direction = "backward", trace = 0, k = log(n))

summary(fit_reduced)

rmse(cars_trn$price, predict(fit_reduced, cars_trn))

rmse(cars_tst$price, predict(fit_reduced, cars_tst))

calc_loocv_rmse(fit_common)

calc_loocv_rmse(fit_reduced)



plot(fitted(fit_reduced), resid(fit_reduced), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)


bptest(fit_reduced)
shapiro.test(resid(fit_reduced))

boxcox(fit_reduced, plotit = TRUE)


fit_cox = lm((price^0.26 - 1)/0.26 ~ engine_displacement + engine_cylinders + 
    highway_fuel_economy + horsepower + maximum_seating + mileage + 
    year + wheel_system + bed_length + franchise_dealer, data = cars_trn)

summary(fit_cox)

bptest(fit_cox)

plot(fitted(fit_cox), resid(fit_cox), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)



```


Reduced model for Pacific region is price ~ engine_displacement + engine_cylinders + 
    highway_fuel_economy + horsepower + maximum_seating + mileage + 
    year + wheel_system + bed_length + franchise_dealer

Predictors reduced to 10, R squared 0.71


We can consider to choose between 0 and 0.3, since the Log-likelihood is almost the same. Apply transformation to price

```{r}

fit_reduced2 = lm(log(price) ~ engine_displacement + engine_cylinders + 
    highway_fuel_economy + horsepower + maximum_seating + mileage + 
    year + wheel_system + bed_length + franchise_dealer, data = cars_trn)

summary(fit_reduced2)

rmse(log(cars_trn$price), predict(fit_reduced2, cars_trn))

rmse(log(cars_tst$price), predict(fit_reduced2, cars_tst))


calc_loocv_rmse(fit_reduced2)


```

From the final model, R squared improves to 0.82, 10 predictors, RMSE is 0.17



```{r}
qqnorm(resid(fit_reduced2), col = "darkgrey")
qqline(resid(fit_reduced2), col = "dodgerblue", lwd = 2)
```


```{r}
pairs(log(price) ~ engine_displacement + engine_cylinders + fuel_tank_volume + highway_fuel_economy + horsepower + maximum_seating + mileage + year + bed_length, data = cars_trn)

fit_col = lm(log(price) ~ engine_displacement + engine_cylinders + 
    highway_fuel_economy + horsepower + maximum_seating + mileage + 
    year + wheel_system + bed_length + franchise_dealer + year:bed_length + year:mileage + maximum_seating:bed_length + 
      engine_cylinders:highway_fuel_economy + engine_cylinders:horsepower + engine_displacement:engine_cylinders +       engine_displacement:highway_fuel_economy, data = cars_trn)

summary(fit_col)

anova(fit_reduced2, fit_col)

rmse(log(cars_trn$price), predict(fit_col, cars_trn))

rmse(log(cars_tst$price), predict(fit_col, cars_tst))


calc_loocv_rmse(fit_col)


```

Adding 7 colinear terms, R squared improved to 0.84, RMSE is 0.16



## Mountain Region

```{r}


set.seed(420420420)

size_num = 12000

cars_idx = sample(nrow(cars_mountain), size_num)

cars_trn = cars_mountain[cars_idx, ]

cars_tst = cars_mountain[-cars_idx, ]


fit_common = lm(price ~ engine_displacement + engine_cylinders + highway_fuel_economy + horsepower + maximum_seating + transmission + 
                  mileage + year + wheel_system + power_hp + bed_length +
                  cabin + salvage + franchise_dealer + state, data = cars_trn)


summary(fit_common)


n = length(resid(fit_common))

# Apple BIC
fit_reduced = step(fit_common, direction = "backward", trace = 0, k = log(n))

summary(fit_reduced)

rmse(cars_trn$price, predict(fit_reduced, cars_trn))

rmse(cars_tst$price, predict(fit_reduced, cars_tst))

calc_loocv_rmse(fit_common)

calc_loocv_rmse(fit_reduced)



```


Reduced model for Mountain region is price ~ highway_fuel_economy + horsepower + maximum_seating + 
    mileage + year + wheel_system + power_hp + bed_length + salvage + 
    franchise_dealer + state
Predictors reduced to 11, R squared 0.72

Apply transformation to price

```{r}

fit_reduced2 = lm(log(price) ~ highway_fuel_economy + horsepower + maximum_seating + 
    mileage + year + wheel_system + power_hp + bed_length + salvage + 
    franchise_dealer + state, data = cars_trn)

summary(fit_reduced2)

rmse(log(cars_trn$price), predict(fit_reduced2, cars_trn))

rmse(log(cars_tst$price), predict(fit_reduced2, cars_tst))

calc_loocv_rmse(fit_reduced2)


```

From this model, R squared improves to 0.85, 10 predictors, RMSE is 0.16


```{r}

pairs(log(price) ~ highway_fuel_economy + horsepower + maximum_seating + mileage + year + power_hp, data = cars_trn)


fit_col = lm(log(price) ~ highway_fuel_economy + horsepower + maximum_seating + 
    mileage + year + wheel_system + power_hp + bed_length + salvage + 
    franchise_dealer + state + year:bed_length + 
    year:mileage + maximum_seating:bed_length + highway_fuel_economy:maximum_seating + horsepower:year, data = cars_trn)


summary(fit_col)

anova(fit_reduced2, fit_col)

rmse(log(cars_trn$price), predict(fit_col, cars_trn))

rmse(log(cars_tst$price), predict(fit_col, cars_tst))


calc_loocv_rmse(fit_col)

```






## Comparison

Comparing the models from the two regions, besides the common predictors, pacific region prefers engine_displacement, engine_cylinders, but doesn't have state as predictor.
Mountain region model prefers power_hp, salvage and state.



