---
title: "botelho3-final-project"
author: "botelho3"
date: "26/06/2021"
output: html_document
---

```{r include=FALSE}
source('includes.R')
source('utils.R')
knitr::opts_chunk$set(echo = TRUE)
set.seed(07042020)
```

```{r include=FALSE}
#cars_df = readr::read_csv("../../used_truck_data_clean.csv")
cars_df = readr::read_csv("../../used_truck_data_with_region.csv")
```

```{r}
spec(cars_df)
total_columns = ncol(cars_df)
west_north_central = c("ND", "MN", "SD", "NE", "IA", "KS", "MO")
west_south_central = c("OK", "AR", "TX", "LA")
my_states = c(west_north_central, west_south_central)
my_df = cars_df[, state=my_states]
my_df$wheel_system = as.factor(my_df$wheel_system)
# To allow log(mileage to work make any mileage < 1 == 1) so log goes from -Inf
# to 0.
my_df[my_df[, "mileage"] < 1.0, "mileage"] = 1.0
unique(cars_df$region)
```

```{r include=FALSE}
free_memory(cars_df)
```

#### Initial Experiments

```{r}
pairs_columns = c("price", "engine_cylinders", "engine_displacement", "mileage", "daysonmarket", "year", "listed_date")
pairs_df = random_rows_from_df(my_df, 0.10)[, pairs_columns]
pairs(pairs_df, col = "dodgerblue")
```

Notes from QQ plot. 
- The relationship of price to mileage looks like it might be better modeled by
1/x. Log(x) would probably work too because of the orders of magnitude.
- engine_cylinders and engine_displacement are collinear. Let's use
engine_displacement since it's an integer not a factor.
- year is almost perfectly linear + quadratic with price.
- daysonmarket is nor very useful lots of noise

```{r}
pairs_columns2 = c("price", "wheel_system", "power_hp", "torque_lb_ft", "maximum_seating")
pairs_df = random_rows_from_df(my_df, 0.10)[, pairs_columns2]
pairs(pairs_df, col = "darkorange")
```

- power_hp and torque_lb_ft are colinear
- maximum_seating is strongly linear with price.
- wheel system is also strongly linear 

```{r}
free_memory(pairs_df)
```


```{r}
fit1 = lm(price ~ year + I(year^2) + engine_displacement + log(mileage) +
   maximum_seating + wheel_system + power_hp + bed_length +
  frame_damaged + salvage,
   data = my_df)
summary(fit1)
do_fitted_vs_residual = function(model = NULL, main = "Fitted vs. Residuals",
                                 remove_outliers = FALSE) {
  std = sd(resid(model))
  if (remove_outliers) {
    keep = resid(model) > -900000 & resid(model) < 900000 
  } else {
    keep = rep(TRUE, length(resid(model)))
  }
  plot(fitted(model)[keep],
       resid(model)[keep], col = "dodgerblue", pch = 20,
       xlab = "Fitted", ylab = "Residuals", main = main)
  abline(h = 0, col = "darkorange", lwd = 2)
  keep
}
do_fitted_vs_residual(fit1, remove_outliers = FALSE)
```



















